Evaluating computational methods for modeling off-normal operation of gas centrifuge cascades

### 2020-12-08 abstract.tex
formatting abstract and title page
reverting italics abstract
proofreading

The goal of this work focuses on developing the necessary understanding of
potential misuse of enrichment cascades, contributing to more effective
international safeguards designs and approaches.  While it is straightforward to
design a symmetric enrichment cascade under ideal conditions as a function of the
theoretical feed, product, and tails assays, it is very difficult to find
reliable information about the behavior of a given cascade when the feed assay
does not match the design value. Several methods have been developed to assess
the behavior of an enrichment cascade in such circumstances. In addition to the
cut ($\theta$) those methods evaluate the feed to product, feed to tails, and the
product to tails enrichment ratio, $\alpha$, $\beta$ and $\gamma$, respectively,
as a function of the cascade feed assay. As those four parameters depend on each
other, determining two of them fully defines the other.  The first approach
consists of fixing $\theta$ and $\alpha$ recomputing the corresponding assays at
each stages of the cascade. The second one maintains the ideal condition of the
cascade ($\alpha$ and $\beta$ fixed across the whole cascade), modifying
$\theta$ value at each stage accordingly. Both approaches have been implemented
into the Cyclus fuel cycle simulator\cite{cyclus, mbmore.2018}. The third fixes
$\theta$ and $\gamma$, using both $\alpha$ and $\beta$ at each stage as free
parameters. The third method has been investigated in \cite{walker.2017}.

>>>>

The goal of this work focuses on developing the necessary understanding of
potential misuse of enrichment cascades, contributing to more effective
international safeguards designs and approaches.  While it is straightforward to
design a symmetric enrichment cascade under ideal conditions as a function of the
theoretical feed, product, and tails assays, it is very difficult to find
reliable information about the behavior of a given cascade when the feed assay
does not match the design value. Several methods have been developed to assess
the behavior of an enrichment cascade in such circumstances. In addition to the
cut ($\theta$) those methods evaluate the feed to product, feed to tails, and the
product to tails enrichment ratio, $\alpha$, $\beta$ and $\gamma$, respectively,
as a function of the cascade feed assay. As those four parameters depend on each
other, determining two of them fully defines the other.  The first approach
consists of fixing $\theta$ and $\alpha$, recomputing the corresponding assays at
each stages of the cascade. The second one maintains the ideal condition of the
cascade ($\alpha$ and $\beta$ fixed across the whole cascade), modifying
$\theta$ values at each stage accordingly. Both approaches have been implemented
into the Cyclus fuel cycle simulator\cite{cyclus, mbmore.2018}. The third fixes
$\theta$ and $\gamma$, using both $\alpha$ and $\beta$ at each stage as free
parameters. The third method has been investigated in \cite{walker.2017}.

### 2020-12-08 dis_clusion.tex
proofreading

We can observe that when the cascade is left completely untouched (Model A) or
when it is slightly retuned to maintain the tails to product enrichment factor
as well as the cut of each centrifuges (Model C), chaining the cascade can
achieve large increase of the enrichment at each level.  On the contrary, when
retuning the cut of each centrifuges to maintain the ideal state of the cascades
(Model B) while chaining them, the \gls{HEU} production rate is favored over the
enrichment gain.

The tails recycling allows each model to achieve a large gain in productivity,
even for then model B in which the number of levels required to reach $90w\%$ of
$^{235}$U in the uranium does not change. Even if no cascade chaining options
achieves the same production rate as direct enrichment, the model B with tail
recycling reached about $80\%$ of the optimum production rate. Such production
rate would allow the accumulation of a Significant Quantity of \gls{HEU} in less
than 8 months\ldots

@@

This work has investigated the possibility to chain centrifuge enrichment
cascades that are designed to enrich uranium for commercial reactors in order to
produce \gls{HEU}, instead. Three methods have been implemented to model enrichment
cascade behavior when fed with different uranium enrichment than the designed
enrichment.

%% This work has investigated and quantified the difference between potential
%% models for retuning of a centrifuge enrichment cascade in order to chain them to produce
%% \gls{HEU} initially tuned to produce uranium enrichment for commercial reactors.
One of these method achieves up to $80\%$ of the production rate of a single
large enrichment cascade designed specifically for \gls{HEU} production using
the same number of centrifuges.

>>>>

We can observe that when the cascade is left completely untouched (Model A) or
when it is slightly retuned to maintain the tails to product enrichment factor
as well as the cut of each centrifuges (Model C), chaining the cascade can
achieve large increase of the enrichment at each level.  On the contrary, when
retuning the cut of each centrifuge to maintain the ideal state of the cascades
(Model B) while chaining them, the \gls{HEU} production rate is favored over the
enrichment gain.

The tails recycling allows each model to achieve a large gain in productivity.
Even if no cascade chaining options achieves the same production rate as direct
enrichment, all models with tail recycling reached close to $73\%$ of the optimum
production rate. Such a production rate would allow the accumulation of a
Significant Quantity of \gls{HEU} in less than 8 months\ldots

@@

This work has investigated the possibility to chain centrifuge enrichment
cascades that are designed to enrich uranium for commercial reactors in order to
produce \gls{HEU} instead. Three methods have been implemented to model
symmetric enrichment cascade behavior when fed with different uranium enrichment
than designed.

%% This work has investigated and quantified the difference between potential
%% models for retuning of a centrifuge enrichment cascade in order to chain them to produce
%% \gls{HEU} initially tuned to produce uranium enrichment for commercial reactors.
Each method achieves up to $73\%$ of the production rate of a single
large enrichment cascade designed specifically for \gls{HEU} production using
the same number of centrifuges.

### 2020-12-08 experiement.tex
scaling cascade config schematic figure
proofreading

\section{The experiment}
This work focuses on comparing the different misuse models to a reference
calculation in which a single large cascade is build and designed to directly
produce \gls{HEU} from natural uranium.
This works uses the Cyclus fuel cycle simulator to allow material exchange
between facilities. The enrichment cascade algorithm have been implemented in
the $mbmore$ package \cite{mbmore.2018}.
In each cases, 5060 centrifuges have been used and spread across up to 30
different gas centrifuge enrichment cascades.

@@

\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.45]{flow}
    \caption{Schematic representation of the chained cascades with three levels,
    with the feed, product and the tails flows, in blue, green and red,
    respectively. The dashed orange line represent the alternative tails flow when
    tails recycling is considered.}
    \label{fig:cascade_level}
\end{figure}

@@

A flow equation similar to \eqref{eq:flow} is then solved to obtain the optimum
number of cascade per level. When the tails are not recycled, the $(1-\theta)$
terms are removed from the flow equation.  The results of the level population
are summarized in Table \ref{tab:level}.

As it is not possible to assign a fraction of an enrichment cascade, cascade per
level are rounded up for each level but the first one. The remaining available
cascades are attributed to the first level.

>>>>

\section{The experiment}
This work focuses on comparing the different misuse models to a reference
calculation in which a single large cascade is designed and built to directly
produce \gls{HEU} from natural uranium.
This works uses the Cyclus fuel cycle simulator to allow material exchange
between facilities. The enrichment cascade algorithm has been implemented in
the $mbmore$ package \cite{mbmore.2018}.
In each cases, 5060 centrifuges have been used and spread across up to 30
different gas centrifuge enrichment cascades. This is inspired by limitations
placed on the Iranian uranium enrichment program via the Joint Comprehensive
Plan of Action (JCPoA) \cite{jcpoa}.

@@

\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[width=\linewidth]{flow}
    \caption{Schematic representation of the chained cascades with three levels,
    with the feed, product and the tails flows, in blue, green and red,
    respectively. The dashed orange line represent the alternative tails flow when
    tails recycling is considered.}
    \label{fig:cascade_level}
\end{figure}

@@

A flow equation similar to \eqref{eq_flow} is then solved to obtain the optimum
number of cascade per level. When the tails are not recycled, the $(1-\theta)$
terms are removed from the flow equation.  The results of the level population
are summarized in Table \ref{tab:level}.

As it is not possible to assign a fraction of an enrichment cascade, cascade per
level must be rounded and assigned for each level. Again, per the JCPoA, 30
cascades are distributed to the appropriate number of levels to achieve an
HEU threshold with the least amount of levels. This optimal assignment is
manually tested by running many different configurations. In generally, each
level besides the first is rounded up from its optimal fractional value, with
the remaining cascades being assigned to the first level.

### 2020-12-08 motivation.tex
proofreading

With the current international tensions regarding enrichment capabilities, this
work aims to measure the effectiveness of a symmetric enrichment cascade when
used outside of its designed scope and quantify the attractiveness of such way
to build up significant quantity of \gls{HEU}.

The present work investigates the performance of an enrichment cycle when chaining
gas enrichment cascades tuned for low enrich uranium production from natural
uranium. As literature on the mater is for obvious reason limited, three
behavior models have been implemented and used to evaluate the response of an
enrich cascade when fed with different assays than the design one. This work
also takes advantage of the Cyclus\cite{cyclus} fuel cycle capabilities to
evaluate the assay values at equilibrium.

>>>>

With the current international tensions regarding enrichment capabilities, this
work aims to measure the effectiveness of a symmetric enrichment cascade when
used outside of its designed scope and quantify the attractiveness of such ways
to build up significant quantity of \gls{HEU}.

The present work investigates the performance of an enrichment cycle when chaining
gas enrichment cascades tuned for low enriched uranium (LEU) production from natural
uranium to instead produce HEU. Literature on the subject is limited due to its
internationally and politically sensitive nature. Three behavior models have been
implemented and used to evaluate the response of an enrichment cascade when fed
with different assays than originally designed. This work also takes advantage
of the Cyclus \cite{cyclus} fuel cycle simulator's capabilities to evaluate the
assay values at equilibrium.

### 2020-12-08 results.tex
updating real cascade numbers in results table
incrementing results figure scaling width
scaling results figure
scaling results figure further
scaling third subfigure in results
trying one more scaling feature in results table
proofreading

\begin{table}[h!]
\footnotesize
\centering
  \caption{Summary of cascades level population.}
\resizebox{\columnwidth}{!}{%
\begin{tabular*}{\textwidth}{c @{\extracolsep{\fill}} cccccccc}
\toprule

Model       &        &                & A/NR     & A/R      & B/NR     & B/R      & C/NR     & C/R       \\
%Recycling   &       &                & No       & Yes      & No       & Yes      & No       & Yes        \\
\midrule                                                                                                 
        &            & Feed $w\%$     & $0.71$   & $1.31$   & $0.71$   & $0.92$   & $0.71$   & $1.30$ \\
Level 0 & Assay      & Product $w\%$  & $3.97$   & $7.27$   & $3.97$   & $5.10$   & $3.97$   & $7.24$ \\
        &            & Tails $w\%$    & $0.28$   & $0.52$   & $0.28$   & $0.36$   & $0.28$   & $0.51$ \\
        & Cascades   & Impl. (Ideal)  & 25(26.5) & 25(26.5) & 25(26.4) & 24(26.3) & 25(26.5) & 25(26.2)  \\
\midrule                                                                                                 
        &            & Feed $w\%$     & $3.97$   & $11.48$  & $3.97$   & $6.29$   & $3.97$   & $11.64$ \\
Level 1 & Assay      & Product $w\%$  & $21.29$  & $53.30$  & $19.33$  & $27.98$  & $21.41$  & $55.12$ \\
        &            & Tails $w\%$    & $1.68$   & $5.93$   & $1.58$   & $2.54$   & $1.66$   & $5.88$ \\
        & Cascades   & Impl. (Ideal)  & 3(3.1)   & 4(3.1)   & 3(3.1)   & 4(3.1)   & 3(3.1)   & 4(3.4)    \\
\midrule                                                                                                 
        &            & Feed $w\%$     & $21.29$  & $53.3$   & $19.33$  & $31.09$  & $21.41$  & $55.12$ \\
Level 2 & Assay      & Product $w\%$  & $76.39$  & $94.70$  & $58.10$  & $72.31$  & $79.21$  & $96.37$ \\
        &            & Tails $w\%$    & $13.98$  & $47.81$  & $8.51$   & $14.91$  & $13.75$  & $49.65$ \\
        & Cascades   & Impl. (Ideal)  & 1(0.4)   & 1(0.4)   & 1(0.4)   & 1(0.5)   & 1(0.4)   & 1(0.4)   \\
\midrule                                                                                                 
        &            & Feed $w\%$     & $76.39$  & N.A.     & $58.10$  & $72.31$  & $79.21$  & N.A.      \\
Level 3 & Assay      & Product $w\%$  & $98.09$  & N.A.     & $88.92$  & $93.79$  & $98.86$  & N.A.      \\
        &            & Tails $w\%$    & $73.51$  & N.A.     & $35.01$  & $50.36$  & $76.60$  & N.A.      \\
        & Cascades   & Impl. (Ideal)  & 1(0.04)  & N.A.     & 1(0.09)  & 1(0.1)   & 1(0.04)  & N.A.      \\
\midrule                                                                                                 
        &            & Feed $w\%$     & N.A.     & N.A.     & $88.92$  & N.A.     & N.A.     & N.A.      \\
Level 4 & Assay      & Product $w\%$  & N.A.     & N.A.     & $97.89$  & N.A.     & N.A.     & N.A.      \\
        &            & Tails $w\%$    & N.A.     & N.A.     & $75.71$  & N.A.     & N.A.     & N.A.      \\
        & Cascades   & Impl. (Ideal)  & N.A.     & N.A.     & 1(0.04)  & N.A.     & N.A.     & N.A.      \\
\bottomrule
\end{tabular*}
}
\label{tab:level}
\end{table}

As illustrated in Figures \ref{fig:assays} and summarized on Table
\ref{tab:level}, the different models don't have the same effect on the
cascade behavior. While models A and C, achieve a quick enrichment gain with
the cascades chaining, 4/23/78/98 and 4/23/82/99, respectively, model B, only
achieves an enrichment gain of 4/21/61/90.  This table also shows the integer
number of cascades implemented (Impl.) at each level in the simulation, as well as the
non-integer number of cascades that would achieve an ideal configuration (Ideal).

\begin{figure}[h!]
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[scale=0.2]{feed_assays}
        \caption{Feed Assays.}
        \label{sfig:feed_assay}
    \end{subfigure}%
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[scale=0.2]{product_assays}
        \caption{Product Assays.}
        \label{sfig:product_assay}
    \end{subfigure}
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[scale=0.2]{tails_assays}
        \caption{Tails Assays.}
        \label{sfig:tails_assay}
    \end{subfigure}
    \caption{Feed (\subref{sfig:feed_assay}), product
        (\subref{sfig:product_assay}) and tails assays (\subref{sfig:tails_assay})
        in $w\%$ of $^{235}$U, per cascade level from 0 to 3 (red, green, orange,
        blue), per model (A/B/C) and without/with tails recycling (NR/R). The
        black dashed line represents the $90 w\%$ enrichment threshold.}
    \label{fig:assays}
\end{figure}

@@

As shown in Figures \ref{fig:assays}, recycling the tails increases the overall product assay at
all the different levels. As the tails assay of a level $n+1$ is always higher than
the product assay of the level $n-1$, recycling the tails of level $n+1$ will
consequently increase the feed assay of level $n$ (see Table \ref{tab:level}).
Moreover, with an increased feed assay, tails and product assays increase as
well, increasing de facto the feed assays of respectively cascade levels $n-1$
and $n+1$, etc.  This effect reduces the number of cascade levels required to reach
\gls{HEU} in case A and C.

@@

As shown in Figure \ref{fig:HEU_rate}, recycling increases the final \gls{HEU}
production rate, from $2$ to almost $20$ kg/y when using models A and C, and from
$17$ to $38$ kg/y with the model B. For the reference calculation where all the
available cascades are used within a single large cascade design for direct
\gls{HEU} production, the \gls{HEU} production rate is slightly over $50$ kg/y.

As models A and C, rely on maintaining the cut values at each stages of the
cascade and share the same number of levels, have the exact same cascade
repartition across the different levels and the same \gls{HEU} production rate.

>>>>

\begin{table}[h!]
\footnotesize
\centering
  \caption{Summary of cascades level population.}
\resizebox{\columnwidth}{!}{%
\begin{tabular*}{\linewidth}{c @{\extracolsep{\fill}} cccccccc}
\toprule

Model       &        &                & A/NR     & A/R      & B/NR     & B/R      & C/NR     & C/R       \\
%Recycling   &       &                & No       & Yes      & No       & Yes      & No       & Yes        \\
\midrule                                                                                                 
        &            & Feed $w\%$     & $0.71$   & $1.31$   & $0.71$   & $0.92$   & $0.71$   & $1.30$ \\
Level 0 & Assay      & Product $w\%$  & $3.97$   & $7.27$   & $3.97$   & $5.10$   & $3.97$   & $7.24$ \\
        &            & Tails $w\%$    & $0.28$   & $0.52$   & $0.28$   & $0.36$   & $0.28$   & $0.51$ \\
        & Cascades   & Impl. (Ideal)  & 25(26.5) & 25(26.5) & 24(26.4) & 25(26.3) & 25(26.5) & 25(26.2)  \\
\midrule                                                                                                 
        &            & Feed $w\%$     & $3.97$   & $11.48$  & $3.97$   & $6.29$   & $3.97$   & $11.64$ \\
Level 1 & Assay      & Product $w\%$  & $21.29$  & $53.30$  & $19.33$  & $27.98$  & $21.41$  & $55.12$ \\
        &            & Tails $w\%$    & $1.68$   & $5.93$   & $1.58$   & $2.54$   & $1.66$   & $5.88$ \\
        & Cascades   & Impl. (Ideal)  & 3(3.1)   & 4(3.1)   & 3(3.1)   & 3(3.1)   & 3(3.1)   & 4(3.4)    \\
\midrule                                                                                                 
        &            & Feed $w\%$     & $21.29$  & $53.3$   & $19.33$  & $31.09$  & $21.41$  & $55.12$ \\
Level 2 & Assay      & Product $w\%$  & $76.39$  & $94.70$  & $58.10$  & $72.31$  & $79.21$  & $96.37$ \\
        &            & Tails $w\%$    & $13.98$  & $47.81$  & $8.51$   & $14.91$  & $13.75$  & $49.65$ \\
        & Cascades   & Impl. (Ideal)  & 1(0.4)   & 1(0.4)   & 1(0.4)   & 1(0.5)   & 1(0.4)   & 1(0.4)   \\
\midrule                                                                                                 
        &            & Feed $w\%$     & $76.39$  & N.A.     & $58.10$  & $72.31$  & $79.21$  & N.A.      \\
Level 3 & Assay      & Product $w\%$  & $98.09$  & N.A.     & $88.92$  & $93.79$  & $98.86$  & N.A.      \\
        &            & Tails $w\%$    & $73.51$  & N.A.     & $35.01$  & $50.36$  & $76.60$  & N.A.      \\
        & Cascades   & Impl. (Ideal)  & 1(0.04)  & N.A.     & 1(0.09)  & 1(0.1)   & 1(0.04)  & N.A.      \\
\midrule                                                                                                 
        &            & Feed $w\%$     & N.A.     & N.A.     & $88.92$  & N.A.     & N.A.     & N.A.      \\
Level 4 & Assay      & Product $w\%$  & N.A.     & N.A.     & $97.89$  & N.A.     & N.A.     & N.A.      \\
        &            & Tails $w\%$    & N.A.     & N.A.     & $75.71$  & N.A.     & N.A.     & N.A.      \\
        & Cascades   & Impl. (Ideal)  & N.A.     & N.A.     & 1(0.04)  & N.A.     & N.A.     & N.A.      \\
\bottomrule
\end{tabular*}
}
\label{tab:level}
\end{table}

As illustrated in Figures \ref{fig:assays} and summarized on Table
\ref{tab:level}, the different models don't have the same effect on the
cascade behavior. While models A and C achieve a quick enrichment gain with
the cascades chaining, 4/21/76/98 and 4/21/79/99, respectively, model B only
achieves an enrichment gain of 4/19/58/89/98, requiring more levels to even
achieve HEU enrichment. This table also shows the integer
number of cascades implemented (Impl.) at each level in the simulation, compared
to the non-integer number of cascades that would achieve an ideal configuration (Ideal).

\begin{figure}[h!]
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{feed_assays}
        \caption{Feed Assays}
        \label{sfig:feed_assay}
    \end{subfigure}%
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{product_assays}
        \caption{Product Assays}
        \label{sfig:product_assay}
    \end{subfigure}
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{tails_assays}
        \caption{Tails Assays}
        \label{sfig:tails_assay}
    \end{subfigure}
    \caption{Feed (\subref{sfig:feed_assay}), product
        (\subref{sfig:product_assay}) and tails assays (\subref{sfig:tails_assay})
        in $w\%$ of $^{235}$U, per cascade level from 0 to 3 (red, green, orange,
        blue), per model (A/B/C) and without/with tails recycling (NR/R). The
        black dashed line represents the $90 w\%$ enrichment threshold.}
    \label{fig:assays}
\end{figure}

@@

As shown in Figures \ref{fig:assays}, recycling the tails increases the overall product assay at
all the different levels, requiring less levels in the configuration to achieve
the desired enrichment. As the tails assay of a level $n+1$ is always higher than
the product assay of the level $n-1$, recycling the tails of level $n+1$ will
consequently increase the feed assay of level $n$ (see Table \ref{tab:level}).
Moreover, with an increased feed assay, tails and product assays increase as
well, increasing de facto the feed assays of respective cascade levels $n-1$
and $n+1$, etc.  This effect reduces the number of cascade levels required to reach
\gls{HEU} in all cases.

@@

As shown in Figure \ref{fig:HEU_rate}, recycling increases the final \gls{HEU}
production rate, from $5$ to almost $45$ kg/y when using models A and C, and from
$12$ to $45$ kg/y with the model B. For the reference calculation, where all the
available cascades are used within a single large cascade design for direct
\gls{HEU} production, the \gls{HEU} production rate is slightly over $60$ kg/y.

As models A and C rely on maintaining the cut values at each stage of the
cascade and share the same number of levels, they have the exact same cascade
repartition across the different levels and the same \gls{HEU} production rate.

### 2020-12-08 theory.tex
attempting to resize matrix
proofreading

In the following section, equations and algorithms used to model the behavior of
a symmetric ideal gas centrifuge cascade from the individual centrifuge property
are described. Details will also be provided to model the centrifuge behavior
when fed with a different feed enrichment than the design one.

@@

The variable $Z_p$ is the rectifier length, or the location of the feed point, has
an optimal axial location as defined by \cite{raetz.phd}:

@@

This optimizes the recitifer length based on the cut, $\theta$, which is an expression
of the fraction of the centrifuge feed that is output as product, and the counter-current
flow, $L/F$. In practice, this value is a design parameter that is updated in the model
during the design of the centrifuge cascade.

The parameters $r_1$ and $r_2$ are the separation radii of the enriched material
(here ${}^{235}\mathrm{U}$ vs. ${}^{238}\mathrm{U}$), $r_1$ being the withdrawal
radius for the lighter isotope and $r_2$ for the heavier isotope. R\"aetz's
two-shell model looks for optimal values between these two, as defined from
the hydrodynamic equations. The radii ratio is optimized using the following relationship:

@@

A symmetric cascade is a cascade where a stage's feed, $F$, with a tail, $T$, of the next
stage and product, $P$, of the previous one. For the cascade feeding stage
an external feed, $F_{ext}$:

@@

From equation \eqref{eq_raetz} and \eqref{eq_alpha_principle} it is possible to
express $\alpha$ as a function of the feed rate $F$, the separative performance
$\delta U(\theta)$ and the cut $\theta$:

@@

Then, from mass conservation, $N = \theta N' + (1-\theta)N''$ and equation
\eqref{eq_product_assay}, it is possible to express $\beta$ as a function of
only the feed abundance, $R$, the cut $\theta$ and $\alpha$:

@@

Finally from equation \eqref{eq_alpha} and \eqref{eq_beta} it is possible to
determine the cut, $\theta$ required to build an ideal cascade:

@@

To construct an ideal stage, a cut ($\theta$) must first be computed.
This is found iteratively, searching for the optimal cut value where $\alpha = \beta$
for a given feed and centrifuge parameters. With a separative power calculated,
and $\alpha$ and $\beta$ value can be determined for an initial cut guess. This
model assumes that the ideal cut for a stage should be between 0.1 and 0.9. The
two enrichment factors are compared, and the higher or lower cut value is chosen
by which pair of factors are closer. A new cut is determined from the chosen
factors, a new separative power is computed, and new $\alpha$ and $\beta$ values
are compared. This process continues until, to some precision, the resulting
enrichment factors are approximately equal. As illustrated in Fig. \ref{fig_a_m_b},
for a given input feed assay, the ideal feeding stage typically has a $\theta$
value between 0.45 and 0.525 when $\alpha = \beta$.

@@

The second step determines the relative flows at each stages, solving the linear
flow equation, \eqref{eq_flow}.
The cascade can then be populated with actual machines until the maximum number
available of machines is reached.

\begin{equation}
\setcounter{MaxMatrixCols}{20}
\begin{bmatrix}
     -1      & 1-\theta_{_{S+1}} & 0                 & ...  & 0            & 0            & 0             & 0             & 0             & ... & 0               & 0  & 0 \\
\theta_{_S}  & -1                & 1-\theta_{_{S+2}} & ...  & 0            & 0            & 0             & 0             & 0             & ... & 0               & 0  & 0 \\
             &                   &                   &      &              &              & ...           &               &               &     &                 &    &   \\
 0           & 0                 & 0                 & ...  & \theta_{_-2} & -1           & 1-\theta_{_0} & 0             & 0             & ... & 0               & 0  & 0 \\
 0           & 0                 & 0                 & ...  & 0            & \theta_{_-1} & -1            & 1-\theta_{_1} & 0             & ... & 0               & 0  & 0 \\
 0           & 0                 & 0                 & ...  & 0            & 0            & \theta_{_0}   & -1            & 1-\theta_{_2} & ... & 0               & 0  & 0 \\
             &                   &                   &      &              &              & ...           &               &               &     &                 &    &   \\
 0           & 0                 & 0                 & ...  & 0            & 0            & 0             & 0             & 0             & ... & \theta_{_{E-2}} & -1 & 1-\theta_{_E} \\
 0           & 0                 & 0                 & ...  & 0            & 0            & 0             & 0             & 0             & ... & 0               & \theta_{_{E-1}} & -1
 \end{bmatrix}
 \times
 \begin{bmatrix}
     F_{_{S}}   \\
     F_{_{S+1}} \\
     \cdots     \\
     F_{_{-1}}  \\
     F_{_{0} }  \\
     F_{_{1} }  \\
     \cdots     \\
     F_{_{E-1}} \\
     F_{_{E}}
 \end{bmatrix}
 =
 \begin{bmatrix}
     0      \\
     0      \\
     \cdots \\
     0      \\
     F      \\
     0      \\
     \cdots \\
     0      \\
     0
\end{bmatrix}
%\caption{caption needed!}
\label{eq_flow}
\end{equation}

@@

The first one assumes that no changes are being made to the cascade, i.e $\delta
U$, $F$ and $\theta$ are fixed across all stages. The second one assumes the cut
value at each stage is retuned to maintain the ideal state of the cascade,
$\alpha$ and $\beta$ remain fixed. The last one, described in \cite{walker.2017}
assumes the tails to product enrichment factor and the cut remain constants
(eq. \eqref{eq_gamma-def}: $\gamma = \alpha\beta$). Models behaviors and
assumptions are summarized in Tab. \ref{tab:models}.

@@

The tuning method A does not re-optimize $\theta_i$ keeping the same flow as the
ideal configuration. From equation \eqref{eq_alpha}, maintaining $\delta U$ and
$F$ while $\theta$ is unchanged implies $\alpha$ remains unchanged as well.
According to equation \eqref{eq_beta}, when $\alpha$ and $\theta$ are fixed, if
the feed assay ($N$) changes, $\beta$ will change accordingly.  This breaks the
ideal status of the cascade, i.e. $N_{i} \neq N'_{i-1} \neq N''_{i+1}$.

In order to compute the proper product and tails assay at each stage, the tails
and the product from the next and the previous stage respectively must be
blended in order to determine the correct stage feed assay. All feed assays are
iteratively updated, blending the proper product and tails, then using the
updated feed assay, the new product and tails assays are recomputed. This
process is repeated until the sum of the square difference in assays is smaller
than $10^{-8}$.  As the cut remains fixed at each stage, the flows do not need
to be recomputed.

>>>>

In the following section, equations and algorithms used to model the behavior of
a symmetric ideal gas centrifuge cascade from individual centrifuge properties
are described. Details will also be provided to model centrifuge behavior
when fed with a different feed enrichment than designed.

@@

The variable $Z_p$ is the rectifier length, or the location of the feed point,
and has an optimal axial location as defined by \cite{raetz.phd}:

@@

This optimizes the recitifer length based on the cut, $\theta$, which is an expression
of the fraction of the centrifuge feed that is output as product, and the counter-current
flow, $L/F$. In practice, this value is a design parameter that is defined in the model
during the design of the centrifuge cascade.

The parameters $r_1$ and $r_2$ are the separation radii of the enriched material
(here ${}^{235}\mathrm{U}$ vs. ${}^{238}\mathrm{U}$), $r_1$ being the withdrawal
radius for the lighter isotope and $r_2$ for the heavier isotope. R\"aetz's
two-shell model looks for optimal values between these two, as described by
the hydrodynamic equations. The radii ratio is optimized using the following relationship:

@@

A symmetric cascade is a cascade where a stage's feed, $F$, with a tail, $T$, of the next
stage and product, $P$, of the previous one. For the cascade feeding stage
and external feed, $F_{ext}$:

@@

From equation \eqref{eq_raetz} and \eqref{eq_alpha_principle} it is possible to
express $\alpha$ as a function of the feed rate $F$, the separative performance
$\delta U$ (a function of $\theta$), and the cut $\theta$:

@@

Then, from mass conservation, $N = \theta N' + (1-\theta)N''$ and equation
\eqref{eq_product_assay}, it is possible to express $\beta$ as a function of
only the feed abundance, $R$, the cut $\theta$, and $\alpha$:

@@

Finally from equation \eqref{eq_alpha} and \eqref{eq_beta} it is possible to
determine the cut, $\theta$, required to build an ideal cascade:

@@

To construct an ideal stage, a cut, $\theta$, must first be computed.
This is found iteratively, searching for the optimal cut value where $\alpha = \beta$
for a given feed and centrifuge parameters. With a separative power calculated,
an $\alpha$ and $\beta$ value can be determined for an initial cut guess. This
model assumes that the ideal cut for a stage should be between 0.1 and 0.9. The
two enrichment factors are compared, and the higher or lower cut value is chosen
by which pair of factors are closer. A new cut is determined from the chosen
factors, a new separative power is computed, and new $\alpha$ and $\beta$ values
are compared. This process continues until, to some precision, the resulting
enrichment factors are approximately equal. As illustrated in Fig. \ref{fig_a_m_b},
for a given input feed assay, the ideal feeding stage typically has a $\theta$
value between 0.45 and 0.525 when $\alpha = \beta$.

@@

The second step determines the relative flows at each stage, solving the linear
flow equation, \eqref{eq_flow}.
The cascade can then be populated with actual machines until the maximum number
of available machines is reached.

\begin{equation}
\resizebox{\linewidth}{!}{%
$\displaystyle
\setcounter{MaxMatrixCols}{20}
\begin{bmatrix}
     -1      & 1-\theta_{_{S+1}} & 0                 & ...  & 0            & 0            & 0             & 0             & 0             & ... & 0               & 0  & 0 \\
\theta_{_S}  & -1                & 1-\theta_{_{S+2}} & ...  & 0            & 0            & 0             & 0             & 0             & ... & 0               & 0  & 0 \\
             &                   &                   &      &              &              & ...           &               &               &     &                 &    &   \\
 0           & 0                 & 0                 & ...  & \theta_{_-2} & -1           & 1-\theta_{_0} & 0             & 0             & ... & 0               & 0  & 0 \\
 0           & 0                 & 0                 & ...  & 0            & \theta_{_-1} & -1            & 1-\theta_{_1} & 0             & ... & 0               & 0  & 0 \\
 0           & 0                 & 0                 & ...  & 0            & 0            & \theta_{_0}   & -1            & 1-\theta_{_2} & ... & 0               & 0  & 0 \\
             &                   &                   &      &              &              & ...           &               &               &     &                 &    &   \\
 0           & 0                 & 0                 & ...  & 0            & 0            & 0             & 0             & 0             & ... & \theta_{_{E-2}} & -1 & 1-\theta_{_E} \\
 0           & 0                 & 0                 & ...  & 0            & 0            & 0             & 0             & 0             & ... & 0               & \theta_{_{E-1}} & -1
 \end{bmatrix}
 \times
 \begin{bmatrix}
     F_{_{S}}   \\
     F_{_{S+1}} \\
     \cdots     \\
     F_{_{-1}}  \\
     F_{_{0} }  \\
     F_{_{1} }  \\
     \cdots     \\
     F_{_{E-1}} \\
     F_{_{E}}
 \end{bmatrix}
 =
 \begin{bmatrix}
     0      \\
     0      \\
     \cdots \\
     0      \\
     F      \\
     0      \\
     \cdots \\
     0      \\
     0
\end{bmatrix}
$}
%\caption{caption needed!}
\label{eq_flow}
\end{equation}

@@

The first one assumes that no changes are being made to the cascade, i.e $\delta
U$, $F$ and $\theta$ are fixed across all stages. The second one assumes the cut
value at each stage is retuned to maintain the ideal state of the cascade,
$\alpha$ and $\beta$ remain fixed. The last one, described in \cite{walker.2017}
assumes the tails to product enrichment factor and the cut remain constants
(eq. \eqref{eq_gamma-def}: $\gamma = \alpha\beta$). Model behaviors and
assumptions are summarized in Tab. \ref{tab:models}.

@@

The tuning method A does not re-optimize $\theta_i$, keeping the same flow as the
ideal configuration. From equation \eqref{eq_alpha}, maintaining $\delta U$ and
$F$ while $\theta$ is unchanged implies $\alpha$ remains unchanged as well.
According to equation \eqref{eq_beta}, when $\alpha$ and $\theta$ are fixed, if
the feed assay ($N$) changes, $\beta$ will change accordingly.  This breaks the
ideal status of the cascade, i.e. $N_{i} \neq N'_{i-1} \neq N''_{i+1}$.

In order to compute the proper product and tails assay at each stage, the tail
and the product from the next and the previous stage respectively must be
blended in order to determine the correct stage feed assay. All feed assays are
iteratively updated, blending the proper product and tails, then using the
updated feed assay, the new product and tails assays are recomputed. This
process is repeated until the sum of the square difference in assays is smaller
than $10^{-8}$.  As the cut remains fixed at each stage, the flows do not need
to be recomputed.
